---
title: "Homework1"
author: "Gilles Nikles, Simon Erni, Vladyslav Gorbunov"
date: "2024-04-06"
output:
  pdf:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Introduction

Kurze Einführung in das Projekt und die Zielsetzungen.

# 2. Preparings

## 2.1 Libraries

Install the libraries first if not available in the environment:

````{r setup, include=FALSE}
# Uncomment the following install-commands if needed
# install.packages("tsibble")
````

Initialise the needed libraries to execute the rmd-file:

````{r message=FALSE}
# Libraries used in Homework1
library(tidyverse)
library(tsibble)
````

## 2.2 Datasets

Import the Datasets:

````{r message=FALSE}
# Libraries used in Homework1
pop <- read_csv("POP.csv")
psavert <- read_csv("PSAVERT.csv")
unemploy <- read_csv("UNEMPLOY.csv")
````

## 2.3 Combine Datasets

````{r message=FALSE}
# Combine Datasets, we use the 'DATE' field as common column
combined_data <- pop %>%
  inner_join(psavert, by = "DATE") %>%
  inner_join(unemploy, by = "DATE")
````

# 3. Tasks

## 3.1 Task 1

Create a tsibble-Object out of the merged data:
```{r load-saving-rate}
# Create a tsibble-Object
tsibble_data <- as_tsibble(combined_data, index = DATE)
```


In the next step, we create timeplots for the Unemployment and Personal Saving Rate
```{r load-saving-rate}
# Timeplot for Unemployment
ggplot(tsibble_data, aes(x = DATE, y = UNEMPLOY)) +
  geom_line() +
  labs(title = "Timeplot for Unemployment", x = "Date", y = "Unemployment")

# Timeplot for Personal Saving Rate
ggplot(tsibble_data, aes(x = DATE, y = PSAVERT)) +
  geom_line() +
  labs(title = "Timeplot for Personal Saving Rate", x = "Date", y = "Personal Saving Rate")

```


## 3.2 Task 2

In order to compare the data available to us with information on the governing party and its election year, we need to import the data. We use the dataset 'countries' from the collection 'topicalsource' from dolthub. The data is available here:
https://www.dolthub.com/repositories/topicalsource/countries/data/main/leaders

```{r load-saving-rate}
# Import the dataset
america <- read_csv("topicalsource_countries_main_leaders.csv", show_col_types = FALSE)
```

In the next step, we do a short clean-up of the data. This includes removing columns that are not necessary to use in this homework as well as only keeping the parties as of Carter's presidency.

```{r load-saving-rate}
# Keep only party, name, term_start and term_end
america_filtered_party <- america %>%
  select(party, name, term_start, term_end)

# Format date to match the style of the combined_data data
america_filtered_date <- america_filtered_party %>%
  mutate(term_start = as.Date(as.character(term_start), format = "%Y%m%d"),
         term_end = as.Date(as.character(term_end), format = "%Y%m%d"))

# Keep only the dates as of Carter's presidency
america_filtered_years <- america_filtered_date %>%
  filter(term_start >= as.Date("1977-01-01")) %>%
  select(party, term_start, term_end)

# Correct the missing date of Biden's presidency
current_date <- Sys.Date()
america_filtered_years <- america_filtered_years %>%
  mutate(term_end = ifelse(is.na(term_end), current_date, term_end))

rm(america_filtered_party, america_filtered_date)
```

```{r load-saving-rate}
# Keep only party, name, term_start and term_end
america_filtered_party <- america %>%
  select(party, name, term_start, term_end)

# Format date to match the style of the combined_data data
america_filtered_date <- america_filtered_party %>%
  mutate(term_start = as.Date(as.character(term_start), format = "%Y%m%d"),
         term_end = as.Date(as.character(term_end), format = "%Y%m%d"))

# Keep only the dates as of Carter's presidency
america_filtered_years <- america_filtered_date %>%
  filter(term_start >= as.Date("1977-01-01")) %>%
  select(party, term_start, term_end)

rm(america_filtered_party, america_filtered_date)
```

-----> Funktioniert noch nicht!! Next we check,

```{r load-saving-rate}
# Funktion, um die Parteizugehörigkeit zu finden
find_party <- function(date, parties) {
  party <- parties$party[which(parties$term_start <= date & parties$term_end >= date)]
  if (length(party) == 1) {
    return(party)
  } else {
    return(NA)  # Gibt NA zurück, wenn keine Partei gefunden wird
  }
}

# Ergänze die 'Party'-Spalte in 'tsibble_data' durch Iterieren über jede Zeile
tsibble_data <- tsibble_data %>%
  rowwise() %>%
  mutate(Party = find_party(DATE, america_filtered_years))

# Entferne die Zeilweise-Gruppierung, da sie nicht mehr benötigt wird
tsibble_data <- tsibble_data %>%
  ungroup()
```




